---
title: "Writing Some Code"
#author: "user"
#date: "10/18/2014"
output: html_document
runtime: shiny
---

## 1 Distribution Functions

**R** provides functions to work with distributions.  
As an example, consider the **Poisson Distribution** (pois)

For each distribution, four functions are provided with
prefix *d*, *p*, *q*, *r*, yielding

* **dpois** - the density of the Poisson distribution (pdf)
* **ppois** - the cumulative Poisson distribution     (cpf)
* **qpois** - the quantile function computes x for a given probability p=P(X<x)
* **rpois** - random samples drawn from a Poisson distribution

Look at the examples from the previous weeks for distributions you have learned about!

```{r, echo=TRUE}
require(graphics)   # load the graphics functions into R
par(mfrow=c(2,1))   # set up a display of 2x1 plots

lambda_T <- 4
x        <- seq( 0, 10, 1)      # values from 0 to 10 with increment 1
cpf      <- ppois(x, lambda_T)  # cpf for each of the x values
pdf      <- dpois(x, lambda_T)  # pdf for each of the x values

x.5      <- qpois(0.5,lambda_T) # smallest x value for which 0.5 <= P(X<=x)

plot(x,cpf,main='CPF', type='s', xlab='x', ylab='F(x)=P(X<=x)')     # plot with steps
abline( v= x.5, col='red')       # add a vertical line at x=x.5
abline( h=0.5, col='red', lty=3) # add a horizontal dashed line at y=.5

plot(x,pdf,main='PDF (delta functions)', type='h', xlab='x', ylab='f(x)=d/dx P(X<=x)') # plot with stems
```

We can select entries from the Poisson distribution at random.
A histogram of these entries will approach the shape of the pdf

```{r, echo=TRUE}
y<-rpois(1000,lambda_T)    # get 1000 values from a Poisson Distribution
hist(y,breaks=80,freq=FALSE,main='PDF estimate')        # draw a histogram (A lot of breaks to show the discrete values)
```

## 2 Loops, Functions, Sliders and Selectors

### 2.1 A simple loop

```{r, echo=TRUE}
for (i in seq(3)) {  # seq(3) is an array with the values 1,2,3
  print(i)
}
```
Note we could have written "1:3" instead of "seq(3)"  
The **seq** function is more general. Type **"help(seq)"** in the rstudio console window for more information

### 2.2 Assign some values to an array

```{r, echo=TRUE}
vals <- numeric(10)  #set up an array of 10 elements
for (i in 1:length(vals)) {  # loop over all the values of the array
  vals[i] <- mean( rpois(20,lambda_T))
}
print(vals)
```

### 2.3 Functions

Repeat the previous example by defining a function.  
We'll generalize it and allow changes to the number of experiments (10)  
and the number of samples (20) used for the computation of the mean

```{r, echo=TRUE}
# The following defines a function called "compute" with 3 input arguments
compute <- function( n, m, lambda_T) {
    vals <- numeric(m)  #set up an array of m elements
    for (i in 1:length(vals)) {  # loop over all the values of the array
      vals[i] <- mean( rpois(n, lambda_T))
    }
    vals  # the value(s) returned by the function on the last line
}

# now call the function
v <- compute(20, 10, 4)
print(v)
```

We can get fancy and pass the distribution function as an argument:  
We will generalize the previous example to allow us to compute the mean of any distribution

```{r, echo=TRUE}
# The following defines a function called "my_func" to compute samples of a distribution
# The intent is to make sure it takes a single argument n
# The global variable "lambda_T" would be visible and could be used directly
#     I chose to pass the value 4 to "rpois" directly
my_func <- function(n) {
    rpois(n, 4)
}

# The new "compute" function now gets an argument "func" that will be called with a single argument
compute_my_func <- function( func, n, m ) {
    vals <- numeric(m)  #set up an array of m elements
    for (i in 1:length(vals)) {  # loop over all the values of the array
      vals[i] <- mean( func(n) )
    }
    vals  # the last line in a function is the result, i.e., the set of output value(s)
}

# now call the function with "my_func" and assign its output values to "v"
v <- compute_my_func( my_func, 20, 10 )
print(v)
```

### 2.4 Display some input controls and a graph

```{r, echo=TRUE}
# First define the interface elements:
#     We are adding a selection box with value input$n_breaks
#     and a slider with value input$bandwidth
inputPanel(
    selectInput("n_breaks", label = "Number of bins:",
                choices = c(10, 20, 35, 50), selected = 20),   # note the comma at the end of this line:
                                                               # we are about to add the slider
  
    sliderInput("bandwidth", label = "Bandwidth adjustment:",
                min = 0.2, max = 2, value = 1, step = 0.2)
)

# we now define the plot
renderPlot({
    #functions, computations and a plot go here
    # note we could reuse the functions we defined above, i.e.,
    print( compute_my_func(my_func,3,4))
  
    # Compute some samples from the normal distributions
    # We'll draw a histogram and something more fancy:
    #     sort of an interpolated form of the histogram (i.e., density)
    y    <- rnorm(50)   # 50 samples from a normal distribution
    dens <- density(y, adjust = input$bandwidth) # weighted average of neighboring samples, where
                                                 # bandwidth determines the number of neighboring samples
 par(mfrow=c(1,2))
  
    hist(y,  probability = TRUE, breaks=as.integer(input$n_breaks)) # draw the histogram
                #  the "as.integer()" function insures breaks is assigned an integer value

    lines(dens, col = "blue") # add the density
    x<-seq(min(y),max(y),.1)
    lines(x,dnorm(x), col="red") # add a plot of the normal

hist(y,  probability = TRUE, breaks=as.integer(input$n_breaks)) # draw the histogram

})
```

Change the number of samples from 50 to 5000 in the example above. What happens?
Can you change the example to make the number of samples another input, e.g., using   
sliderInput("num", label = "Number of samples:", min = 10, max = 10000, value = 50, step = 10)

## 3 Optional Homework Assignment

Write an rstudio example like this one, and submit it electronically.  
I want to make your efforts available to the class as a whole!  
**Ask me for help whenever you are stuck!**

* Pick a **distribution** of your choice

* Pick two numbers **N > 0** and **M > 0**

* Experiment: draw **N** samples from the distribution and compute their mean

* Repeat the experiment **M** times, and display a histogram of the **mean** values

* What happens when you increase **N**?

* What happens when you increase the number of values **M**?

* Repeat the experiment with a computation of the standard deviation instead of the mean

* What happens if you choose a different distribution?

** Basically, you only need to modify one or two lines  
as you stitch together the examples above. **
If you want to break the text and the code into separate files,
look at what I did with the Distribution Examples I gave you previously.
