---
title: "Chapter2"
output: html_document
runtime: shiny
---

# Basic Probability Computations

## The prob package

A useful package is called **'prob'**
* Type **help.start()** in the console window,
* select **"Packages"** followed by
* **"User guides, package vignettes and other documentation"**, and
* look at the **prob::prob** PDF file.

Include it with the library command: **library(prob)**
```{r, echo=FALSE, message=FALSE}
library(prob)
```

### Highlights - *samples*

Selection from a given set of elements
can occur
* with or without replacement
  + with replacement: an item can be picked any number of times,
    the associated probability does not change
  + without replacement: a gven item can be picked at most once.
    probabilities change after every selection
* The order of the picked items can or cannot be distinguished

The possible results of such a selection process can be
displayed using **urnsamples**, e.g.,
```{r}
urnsamples( c("red","green","blue"), 2, ordered = T, replace=FALSE )
```

**prob** has a number of sample spaces built in,
such as **cards()** and **tosscoin()**

### Highlights - *probabilites of samples*

When the sample space consists of a finite number **N** of samples,
we may be able to attach probabilities to individual samples.
If we assume each sample to be equally likely, we will assign
$P({a}) = \frac{1}{N}$ to sample $a$. Other assumptions are
possible, however: if a coin is not fair, the probability of
a head turning up in a coin toss be me far larger than the
probability of a tail.

With the equiprobable model, the probability of an event $A$
is given by
$$P(A) = \frac{\text{number of samples in A}}{N},$$
so that computing probabilities reduces to a counting exercise.

The number of possible results of the selection process from an urn is computed by **nsamp**,
with arguments *nitems*, *k elections*, *replace* and *ordered*.
```{r, echo=TRUE}
nsamp( 3, 2, ordered = T, replace=FALSE )
```
Note the arguments of **nsamp** can be vectors.

Frequencies associated with samples can of course be computed directly, e.g., by using the **factorial()** and **choose()** functions. Powers are computed with the notation *n^k*
```{r,echo=TRUE}
c(2^5,factorial(5),choose(20,3))
```

For equally likely outcomes
we can use the *makespace=TRUE* argument to attach a *probs* column to the samples
```{r, echo=TRUE}
tosscoin(2,makespace=T)
```

Events are subsets of the sample space. In the simple case we
are discussing here, the probability of
a given event is the sum of the probabilities of the individual samples.
```{r, echo=TRUE}
M <- tosscoin(3,makespace=TRUE)
A <- subset( M, M$toss2=="T")
M
A
sum(A$probs)
```
