---
title: "Discrete Distributions"
#author: "user"
#date: "09/26/2014"
output: html_document
runtime: shiny
---
```{r,echo=FALSE}
source('distributions.r')
```

# 1 - Selection without replacement
## The Hypergeometric Distribution
* **N** objects, of which **K** are classified as successes
* **n** objects are selected at random without replacement, so $p=\frac{K}{N}$

Canonical problem: sampling without replacement

*When $N$ is large and $n$ is small, this distribution can be approximated by the Binomial Distribution*

**X** determines the number of successes in the selection

$$f(x) = \frac{C^K_x C^{N-K}_{n-x}}{C^N_n}, \quad E(X) = n p, \quad Var(X)= n p (1-p) \frac{N-n}{N-1}$$
```{r, echo=FALSE}
inputPanel(
  sliderInput("n1_hyper", label = "Available Successes",
              min = 1, max = 500, value = 20, step = 1)
  ,sliderInput("n2_hyper", label = "Available Failures",
              min = 1, max = 500, value = 80, step = 1)
  ,sliderInput("n_hyper", label = "Number of items selected (can't exceed N)",
              min = 1, max = 1000, value = 10, step = 1)
  ,checkboxInput( "binom_nhyper", label="Show Binomial",value=FALSE)
)

renderPlot({
  plothyper( input$n1_hyper, input$n2_hyper, input$n_hyper, input$binom_nhyper )
})
```

# 2 - Bernoulli Trials
## The Binomial Distribution
* **N** *independent* Bernoulli trials (success/failure)
* **p** constant probability of success

Canonical Problem:  sampling with replacement; coin toss   
Important Property: memoryless

**X** determines the number of successes in **N** trials

$$f(x) = C^N_x p^x (1-p)^{N-x}, \quad E(X) = N p, \quad Var(X)= N p (1-p)$$
```{r, echo=FALSE}
inputPanel(
  sliderInput("num_binom", label = "Number of Trials N:",
              min = 1, max = 500, value = 20, step = 1)
  ,sliderInput("p_binom", label = "Probability p",
              min = 0.01, max = 1, value = 0.5, step = .01)
  ,checkboxInput( "binom_normal", label="Show Normal",value=FALSE)
)

renderPlot({
  plotbinom( input$num_binom, input$p_binom, input$binom_normal)
})
```

## The Geometric Distribution
* **N** *independent* Bernoulli trials (success/failure)
* **p** constant probability of success 

Canonical Problem:  sampling with replacement; coin toss   

**X** determines the number of trials until a success is obtained   
      (i.e., length of a run)

$$f(x) = (1-p)^{x-1} p, \quad E(X) = \frac{1}{p}, \quad Var(X)= \frac{1-p}{p^2}$$
```{r, echo=FALSE}
inputPanel(
  sliderInput("n_geom", label = "Max x value to be plotted",
              min = 1, max = 500, value = 10, step = 1)
  ,sliderInput("p_geom", label = "Probability p",
              min = 0.01, max = 1, value = 0.5, step = .01)
)

renderPlot({
  plotgeom( as.numeric(input$p_geom), input$n_geom )
})
```

## The Negative Binomial Distribution
* **N** *independent* Bernoulli trials (success/failure)
* **p** constant probability of success 

**X** determines the number of trials until **r** successes are obtained

$$f(x) = C^{x-1}_{r-1}(1-p)^{x-r} p^r, \quad E(X) = r \frac{1}{p}, \quad Var(X)= r \frac{1-p}{p^2}$$
```{r, echo=FALSE}
inputPanel(
  sliderInput("n_nbinom", label = "Max x value to be plotted",
              min = 1, max = 500, value = 10, step = 1)
  ,sliderInput("r_nbinom", label = "Number of required successes r",
              min = 1, max = 500, value = 2, step = 1)
  ,sliderInput("p_nbinom", label = "Probability p",
              min = 0.01, max = 1, value = 0.5, step = .01)
)

renderPlot({
  plotnbinom( input$r_nbinom, input$n_nbinom, input$p_nbinom )
})
```

# 3 - Events with Constant Ave Rate of Occurrence
## The Poisson Distribution

A random experiment results in events occurring over an interval $T$.  
Derived from the Binomial Distribution by modeling events in a small subinterval as the successful outcome of a Bernoulli trial, and only recording the subinterval of the event 

Let **X_t** be the number of events in an interval of length $t$,
and consider subintervals $\Delta T$ of  $T$.

If the results are such that as $\Delta T \rightarrow 0$

* $P( X_{\Delta T} > 1) \rightarrow 0$
* $P( X_{\Delta T} = 1) \rightarrow \lambda \Delta T$ for some constant $\lambda > 0$
* **an event in any subinterval is independent of events in other subintervals**

then $X_T$ has distribution
$$f(x) = \frac{{( \lambda T )}^x}{x !} e^{-\lambda T}, \quad E(X) = \lambda T, \quad Var(X)= \lambda T$$

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_poisson", label = "Number of Points:",
              min = 1, max = 500, value = 20, step = 1)
  ,sliderInput("lambda", label = "lambda T",
              min = 0., max = 20, value = 1., step = .001)
  ,checkboxInput( "poisson_normal", label="Show Normal",value=FALSE)
)

renderPlot({
  plotpoisson( input$lambda, extent=input$n_poisson, show_normal=input$poisson_normal )
})
```

